### 2012-02-11 trunk/GUIErrorDetector/paper/evaluation.tex
Spacing hacks

\begin{myindentpar}{-2mm}
\begin{itemize}%\addtolength{\itemsep}{-0.5\baselineskip}
\Item \textbf{SWT desktop applications.}
%VirgoFTP~\cite{virgo} and FileBunker~\cite{filebunker} are two useful SWT desktop applications.
VirgoFTP~\cite{virgo} implements a simple FTP client based on Java and the SWT UI library.
FileBunker~\cite{filebunker} is a file backup application that uses one or more GMail
accounts as its backup repository. ArecaBackup~\cite{areca} offers a local
file backup solution for Linux and Windows.

\vspace{0.5mm}
\vspace{0.5mm}

\Item \textbf{Eclipse plugins.}
HudsonEclipse~\cite{hudson} monitors Hudson build status from Eclipse.
EclipseRunner~\cite{eclipserunner} extends Eclipse's capability of running launch configurations.

\vspace{0.5mm}

\Item \textbf{Swing applications.} %S3dropbox~\cite{s3dropbox} and SudokuSolver~\cite{sudokusolver} are two Swing desktop applications.
S3dropbox~\cite{s3dropbox}  allows users
to drag and drop files to their Amazon S3 accounts. SudokuSolver~\cite{sudokusolver}
computes Sudoku solutions using mutlithreaded execution.

\vspace{0.5mm}

\Item  \textbf{Android applications.} %MyTracks~\cite{mytracks}, Fennec~\cite{fennec}, and SGTPuzzles~\cite{sgtpuzzles} are three Android applications.
SGTPuzzler~\cite{sgtpuzzles} is a single-player logic game.
Fennec~\cite{fennec}, developed by Mozilla, is the Mozilla
Firefox web browser for mobile devices. 
MyTracks~\cite{mytracks}, developed by Google, records users' GPS tracks, and provides
interfaces to visualize them on Google Maps.

\end{itemize}
\end{myindentpar}

@@

We first inspected the source code of each subject to determine whether any
native methods or project-specific GUI interaction patterns were implemented.
Only SGTPuzzler uses native methods to interact with
the underlying operating system. For it, we manually checked the possible
Java methods that a native method may call, and then added \annotationnum \CodeIn{@CalledByNativeMethods}
annotations for it. Two subjects, MyTracks and Fennec, employ a customized pattern
to interact with the GUI framework. For example, in MyTracks, all non-UI
threads are initialized via using the library method \CodeIn{android.os.} \CodeIn{handler.handleCallback}.
% the \CodeIn{safeInvokeXX}
%method in the XXX subject first checks whether the current thread is the UI-thread, and
%then determines whether to direct access a GUI object or wrap the code with
%message passing. 
We added \filternum user-defined filters for these two subjects.
%to indicate
%that they are safe to be invoked from any thread. 
In this experiment, none of the paper
authors was familiar with the subjects, but we found it was quite easy
to add extra annotations and user-defined filters. All these
manual parts took less than a total of 60 minutes.

@@

Our error detection algorithm in Figure~\ref{fig:detectalgorithm} uses a
slightly modified BFS to search for
potential single-GUI-thread violations. However, standard BFS
and other graph search strategies such as
DFS and exhaustive path search can also be employed. To
understand trade-offs between using different search strategies, we 
implemented three variants of the algorithm in Figure~\ref{fig:detectalgorithm}, and
evaluated them on the same subjects with the identical environment settings.

The first variant uses standard BFS for graph traversal by deleting
lines 6 and 24 in Figure~\ref{fig:detectalgorithm}, and changing
7 in Figure~\ref{fig:detectalgorithm} to:

\vspace{0.2mm}
$\mathit{worklist}$ $\leftarrow$ $\bigcup_{entry \in entryNodes}$ getReachableStarts($\mathit{entry}$)
\vspace{0.2mm}

\noindent This variant returns the shortest path from \textit{any} \CodeIn{Thread.start()} node
to \textit{any} UI-accessing node, rather than the shortest path from \textit{a given}
\CodeIn{Thread.start()} node to \textit{any} UI-accessing node as the algorithm in
Figure~\ref{fig:detectalgorithm} does.
This variant found 10 errors in total and missed 1 error in S3dropbox and 1 error in
SudokuSolver. This is because in BFS (either the standard BFS or our algorithm), each
node will \textit{only} be visited once; but in the standard BFS, if a node has been covered
in an error path from a \CodeIn{Thread.start()} node, that node will not be visited again
even when the standard BFS traverses from another \CodeIn{Thread.start()} node, thus
missing potential errors.

The second variant uses DFS for graph traversal by changing
the data structure of $\mathit{worklist}$ on line 7 in Figure~\ref{fig:detectalgorithm} to
a stack, and changing queue operations (\CodeIn{dequeue} and \CodeIn{enqueueAll}
on lines 10 and 21 in Figure~\ref{fig:detectalgorithm}, respectively)
to the corresponding stack operations (\CodeIn{pop} and \CodeIn{pushAll}).
This variant found 11 errors in total and missed 1 error in the FileBunker subject.
The reason is that DFS tends to search deeper into the graph, and the path it returns
may not be the shortest one and is more likely to be an infeasible path in practice.
%For the FileBunker subject, paths
%returned by DFS are much longer than BFS, and do not actually exist.

The third variant uses exhaustive search to find potential errors. This variant
enumerates all non-cyclic paths from all reachable \CodeIn{Thread.start()} nodes
to each UI-accessing node, and then checks whether each
path spawns a new thread and accesses GUI objects without using safe UI methods.
This variant ran out of memory on all subjects. We checked
the explored paths before the algorithm terminated, and found none of them
revealed a real error. This is because the number of
non-cyclic paths in a graph is exponential in the graph size.
For a realistic call graph which often contains thousands of nodes,
it is infeasible to enumerate all paths.
%only a very few amount of paths can be enumerated within a practical amount of time.

>>>>

\vspace{-5pt}

% \begin{myindentpar}{-2mm}
\begin{itemize}%\addtolength{\itemsep}{-0.5\baselineskip}
\item \textbf{SWT desktop applications.}
%VirgoFTP~\cite{virgo} and FileBunker~\cite{filebunker} are two useful SWT desktop applications.
VirgoFTP~\cite{virgo} implements a simple FTP client based on Java and the SWT UI library.
FileBunker~\cite{filebunker} is a file backup application that uses one or more GMail
accounts as its backup repository. ArecaBackup~\cite{areca} offers a local
file backup solution for Linux and Windows.

% \vspace{0.5mm}
% \vspace{0.5mm}

\item \textbf{Eclipse plugins.}
HudsonEclipse~\cite{hudson} monitors Hudson build status from Eclipse.
EclipseRunner~\cite{eclipserunner} extends Eclipse's capability of running launch configurations.

% \vspace{0.5mm}

\item \textbf{Swing applications.} %S3dropbox~\cite{s3dropbox} and SudokuSolver~\cite{sudokusolver} are two Swing desktop applications.
S3dropbox~\cite{s3dropbox}  allows users
to drag and drop files to their Amazon S3 accounts. SudokuSolver~\cite{sudokusolver}
computes Sudoku solutions using mutlithreaded execution.

% \vspace{0.5mm}

\item  \textbf{Android applications.} %MyTracks~\cite{mytracks}, Fennec~\cite{fennec}, and SGTPuzzles~\cite{sgtpuzzles} are three Android applications.
SGTPuzzler~\cite{sgtpuzzles} is a single-player logic game.
Fennec~\cite{fennec}, developed by Mozilla, is the Mozilla
Firefox web browser for mobile devices. 
MyTracks~\cite{mytracks}, developed by Google, records users' GPS tracks, and provides
interfaces to visualize them on Google Maps.

\end{itemize}
% \end{myindentpar}

@@

We first inspected the source code of each subject to determine whether any
native methods or project-specific GUI interaction patterns were implemented.
Only SGTPuzzler uses native methods to interact with
the underlying operating system. For it, we manually checked the possible
Java methods that a native method may call, and then added \annotationnum \CodeIn{@CalledByNativeMethods}
annotations for it. Two subjects, MyTracks and Fennec, employ a customized pattern
to interact with the GUI framework. For example, in MyTracks, all non-UI
threads are initialized via using the library method \CodeIn{android.os.}\discretionary{}{}{}\CodeIn{handler.handleCallback}.
% the \CodeIn{safeInvokeXX}
%method in the XXX subject first checks whether the current thread is the UI-thread, and
%then determines whether to direct access a GUI object or wrap the code with
%message passing. 
We added \filternum user-defined filters for these two subjects.
%to indicate
%that they are safe to be invoked from any thread. 
In this experiment, none of the paper
authors was familiar with the subjects, but we found it was quite easy
to add extra annotations and user-defined filters. All these
manual parts took less than a total of 60 minutes.

@@

The error detection algorithm in Figure~\ref{fig:detectalgorithm} uses a
separate BFS for each entry node.  This section evaluates three variants:
a multiple-source BFS; a separate DFS for each entry node; and exhaustive
search.

The first variant uses a single BFS, but starting at multiple sources.  It deletes
lines 6 and 24 in Figure~\ref{fig:detectalgorithm} and changes line
7 to:

\vspace{0.2mm}
$\mathit{worklist}$ $\leftarrow$ $\bigcup_{\mathit{entry} \in \mathit{entryNodes}}$ getReachableStarts($\mathit{entry}$)
\vspace{0.2mm}

\noindent This variant returns the shortest path from any \CodeIn{Thread.start()} node
to any UI-accessing node, rather than the shortest path from each
\CodeIn{Thread.start()} node to any UI-accessing node as the algorithm in
Figure~\ref{fig:detectalgorithm} does.
This variant reported 10 errors and 12 false positives.   It missed 1 error in S3dropbox and 1 error in
SudokuSolver.
It visits every node in the call graph only once, rather than potentially
once per entry node, but in so doing prunes out real error paths.

The second variant one DFS per entry node, by changing
the $\mathit{worklist}$ on line 7 of Figure~\ref{fig:detectalgorithm} to
a stack, and changing queue operations dequeue and enqueueAll
on lines 10 and 21
to pop and pushAll.
This variant reported 11 errors and 10 false positives.  It missed 1 error in FileBunker.
DFS tends to search deeper into the graph and return longer paths that are
more likely to be infeasible or to be removed by the filters.
%For the FileBunker subject, paths
%returned by DFS are much longer than BFS, and do not actually exist.

The third variant uses exhaustive search to find potential errors. This variant
enumerates all non-cyclic paths from all reachable \CodeIn{Thread.start()} nodes
to each UI-accessing node, and then checks whether each
path spawns a new thread and accesses GUI objects without using safe UI methods.
This variant ran out of memory on all subjects, and did not output any real
errors before it terminated.  The number of
non-cyclic paths in a graph is exponential in the graph size, and it 
it is infeasible to enumerate all paths.
for a realistic call graph, which often contains thousands of nodes.
%only a very few amount of paths can be enumerated within a practical amount of time.

Given a sound call graph, exhaustive search is sound because it does not
miss any possible errors.  However, it is impractical.  Compared to
exhaustive search, using BFS or DFS can be viewed as a heuristic filtering
step, akin to the filters of Section~\ref{sec:heuristic}.
%   The algorithm of
% Figure~\ref{fig:detectalgorithm}, which does multiple BFS traversals, does
% less pruning, and runs more slowly, than the variant that does one
% multi-source BFS, but
 The heuristics of Section~\ref{sec:heuristic} are
more effective at retaining real errors than the search variants.
